---
title: "Phase 1 Analysis"
output: html_notebook
---

# adding required libraries

```{r}
library(plyr) #for data manipulation
library(dplyr) #for data manipulation        
library(qcc) #for control charts
library(MSQC) #for maltivariate control charts
library(FactoMineR) #for data analysis
library(factoextra) #for data analysis
library(xlsx) #for reading excel file
library(ggplot2) #for visualization
library(ggbiplot) #for visualization
```

# importing data
```{r}
df <- read.xlsx("C:/Data/isen-614/Project_dataset.xlsx", header=F, sheetIndex = 1)
head(df)
```

## adding column names 
```{r}
for(i in 1:209){
  names(df)[i] <- paste("Col", i)
}

head(df, 10)
```

# Principal Component Analysis

## centering the data

```{r}
df.centered <- scale(df, scale = F, center = T)
```

## finding out Principal components

```{r}
pca_data <- prcomp(df.centered)
summary(pca_data)
```

## Top 10 eigen Values and associated explained variance

```{r}
eig.val <- get_eigenvalue(pca_data)
head(eig.val, 10)
```

# eigen Vectors of top 10 pricipal components

```{r}
pca_data$rotation[1:10, 1:10]
```

## Scree plot

```{r}
fviz_eig(pca_data, addlabels = T, ncp = 15)
```

## Pareto Chart for PCA

```{r}
PCA <- pca_data$sdev[1:10]^2
names(PCA) <- paste("PC", c(1:10))
pareto.chart(PCA)
```

## Plotting biplot and all variables

```{r}
ggbiplot(pca_data)
```

## Data Reduction

```{r}
# considering only 4 PCs which explain more than 80% of variance

df.pca <- pca_data$x[, 1:4]
df.pca <- as.data.frame(df.pca)
dim(df.pca)
```

```{r}
head(df.pca)
```

```{r}
summary(df.pca)
```

## Visualizing Principal Components

```{r}
# PC 1
bw1 <- 2 * IQR(df.pca$PC1) / length(df.pca$PC1)^(1/3)
ggplot(df.pca, aes(x=PC1)) + geom_histogram(aes(y=..density..), binwidth = bw1, colour="black", fill="white") + geom_density(alpha=.2, fill="#FF6666")

# PC 2
bw2 <- 2 * IQR(df.pca$PC2) / length(df.pca$PC2)^(1/3)
ggplot(df.pca, aes(x=PC2)) + geom_histogram(aes(y=..density..), binwidth = bw2, colour="black", fill="white") + geom_density(alpha=.2, fill="#FF6666")

# PC 3
bw3 <- 2 * IQR(df.pca$PC3) / length(df.pca$PC3)^(1/3)
ggplot(df.pca, aes(x=PC3)) + geom_histogram(aes(y=..density..), binwidth = bw3, colour="black", fill="white") + geom_density(alpha=.2, fill="#FF6666")

# PC 4
bw4 <- 2 * IQR(df.pca$PC4) / length(df.pca$PC4)^(1/3)
ggplot(df.pca, aes(x=PC4)) + geom_histogram(aes(y=..density..), binwidth = bw4, colour="black", fill="white") + geom_density(alpha=.2, fill="#FF6666")
```


## X-bar charts of top 4 Principal Components

```{r}
qcc(df.pca[,1], type = "xbar.one", data.name = "PC1")
qcc(df.pca[,2], type = "xbar.one", data.name = "PC2")
qcc(df.pca[,3], type = "xbar.one", data.name = "PC3")
qcc(df.pca[,4], type = "xbar.one", data.name = "PC4")
```

# Hotelling T2 charts for Phase 1 Analysis: 1st Iteration

## Cycle : 1

```{r}
T2.1.1 <- mult.chart(df.pca, type = "t2", alpha=0.0027, phase = 1, method = "sw")
```

```{r}
# out of control points

print(paste("Total out of control data points : ", length(which(T2.1.1$t2 > T2.1.1$ucl)))) 
print(paste("Out of control data points: ", list(which(T2.1.1$t2 > T2.1.1$ucl))))

# removing out of control points
df.pca <- df.pca[-which(T2.1.1$t2 > T2.1.1$ucl),]
print(paste("Total data points after removing out of control data points: ", dim(df.pca)[1]))
```



## Cycle : 2

```{r}
T2.1.2 <- mult.chart(df.pca, type = "t2", alpha=0.0027, phase = 1, method = "sw")
```

```{r}
# out of control points

print(paste("Total out of control data points : ", length(which(T2.1.2$t2 > T2.1.2$ucl)))) 
print(paste("Out of control data points: ", list(which(T2.1.2$t2 > T2.1.2$ucl))))

# removing out of control points
df.pca <- df.pca[-which(T2.1.2$t2 > T2.1.2$ucl),]
print(paste("Total data points after removing out of control data points: ", dim(df.pca)[1]))
```




## Cycle : 3

```{r}
T2.1.3 <- mult.chart(df.pca, type = "t2", alpha=0.0027, phase = 1, method = "sw")
```

```{r}
# out of control points

print(paste("Total out of control data points : ", length(which(T2.1.3$t2 > T2.1.3$ucl)))) 
print(paste("Out of control data points: ", list(which(T2.1.3$t2 > T2.1.3$ucl))))

# removing out of control points
df.pca <- df.pca[-which(T2.1.3$t2 > T2.1.3$ucl),]
print(paste("Total data points after removing out of control data points: ", dim(df.pca)[1]))
```


## Cycle : 4

```{r}
T2.1.4 <- mult.chart(df.pca, type = "t2", alpha=0.0027, phase = 1, method = "sw")
```

```{r}
# out of control points

print(paste("Total out of control data points : ", length(which(T2.1.4$t2 > T2.1.4$ucl)))) 
print(paste("Out of control data points: ", list(which(T2.1.4$t2 > T2.1.4$ucl))))

# total data points

print(paste("Total remaining data points", dim(df.pca)[1]))
```

# m-CUSUM charts for Phase 1 Analysis: 1st Iteration

## Cycle: 1
```{r}
mcusum.1.1 <- mult.chart(df.pca, type = "mcusum2", alpha=0.0027, k = 1.5, h = 6, method = "sw")
```

```{r}
# out of control points

print(paste("Total out of control data points : ", length(which(mcusum.1.1$t2 > mcusum.1.1$ucl)))) 
print(paste("Out of control data points: ", list(which(mcusum.1.1$t2 > mcusum.1.1$ucl))))

# removing out of control points
df.pca <- df.pca[-which(mcusum.1.1$t2 > mcusum.1.1$ucl),]
print(paste("Total data points after removing out of control data points: ", dim(df.pca)[1]))
```

## Cycle: 2

```{r}
mcusum.1.2 <- mult.chart(df.pca, type = "mcusum2", alpha=0.0027, k = 1.5, h = 6, method = "sw")
```

```{r}
# out of control points

print(paste("Total out of control data points : ", length(which(mcusum.1.2$t2 > mcusum.1.2$ucl)))) 
print(paste("Out of control data points: ", list(which(mcusum.1.2$t2 > mcusum.1.2$ucl))))

# removing out of control points
df.pca <- df.pca[-which(mcusum.1.2$t2 > mcusum.1.2$ucl),]
print(paste("Total data points after removing out of control data points: ", dim(df.pca)[1]))
```

## Cycle: 3

```{r}
mcusum.1.3 <- mult.chart(df.pca, type = "mcusum2", alpha=0.0027, k = 1.5, h = 6, method = "sw")
```

```{r}
# out of control points

print(paste("Total out of control data points : ", length(which(mcusum.1.3$t2 > mcusum.1.3$ucl)))) 
print(paste("Out of control data points: ", list(which(mcusum.1.3$t2 > mcusum.1.3$ucl))))

# removing out of control points
df.pca <- df.pca[-which(mcusum.1.3$t2 > mcusum.1.3$ucl),]
print(paste("Total data points after removing out of control data points: ", dim(df.pca)[1]))
```

## Cycle: 4

```{r}
mcusum.1.4 <- mult.chart(df.pca, type = "mcusum2", alpha=0.0027, k = 1.5, h = 6, method = "sw")
```

```{r}
# out of control points

print(paste("Total out of control data points : ", length(which(mcusum.1.4$t2 > mcusum.1.4$ucl)))) 
print(paste("Out of control data points: ", list(which(mcusum.1.4$t2 > mcusum.1.4$ucl))))

# removing out of control points
df.pca <- df.pca[-which(mcusum.1.4$t2 > mcusum.1.4$ucl),]
print(paste("Total data points after removing out of control data points: ", dim(df.pca)[1]))
```

## Cycle: 5

```{r}
mcusum.1.5 <- mult.chart(df.pca, type = "mcusum2", alpha=0.0027, k = 1.5, h = 6, method = "sw")
```

```{r}
# out of control points

print(paste("Total out of control data points : ", length(which(mcusum.1.4$t2 > mcusum.1.5$ucl)))) 
print(paste("Out of control data points: ", list(which(mcusum.1.5$t2 > mcusum.1.5$ucl))))

# removing out of control points
df.pca <- df.pca[-which(mcusum.1.5$t2 > mcusum.1.5$ucl),]
print(paste("Total data points after removing out of control data points: ", dim(df.pca)[1]))
```

## Cycle: 6

```{r}
mcusum.1.6 <- mult.chart(df.pca, type = "mcusum2", alpha=0.0027, k = 1.5, h = 6, method = "sw")
```

```{r}
# out of control points

print(paste("Total out of control data points : ", length(which(mcusum.1.6$t2 > mcusum.1.6$ucl)))) 
print(paste("Out of control data points: ", list(which(mcusum.1.6$t2 > mcusum.1.6$ucl))))

# total data points

print(paste("Total remaining data points", dim(df.pca)[1]))
```

# Hotelling T2 charts for Phase 1 Analysis: 2nd Iteration

## Cycle : 1

```{r}
T2.2.1 <- mult.chart(df.pca, type = "t2", alpha=0.0027, phase = 1, method = "sw")
```

```{r}
# out of control points

print(paste("Total out of control data points : ", length(which(T2.2.1$t2 > T2.2.1$ucl)))) 
print(paste("Out of control data points: ", list(which(T2.2.1$t2 > T2.2.1$ucl))))

# removing out of control points
df.pca <- df.pca[-which(T2.2.1$t2 > T2.2.1$ucl),]
print(paste("Total data points after removing out of control data points: ", dim(df.pca)[1]))
```


## Cycle : 2

```{r}
T2.2.2 <- mult.chart(df.pca, type = "t2", alpha=0.0027, phase = 1, method = "sw")
```

```{r}
# out of control points

print(paste("Total out of control data points : ", length(which(T2.2.2$t2 > T2.2.2$ucl)))) 
print(paste("Out of control data points: ", list(which(T2.2.2$t2 > T2.2.2$ucl))))

# removing out of control points
df.pca <- df.pca[-which(T2.2.2$t2 > T2.2.2$ucl),]
print(paste("Total data points after removing out of control data points: ", dim(df.pca)[1]))
```


## Cycle : 3

```{r}
T2.2.3 <- mult.chart(df.pca, type = "t2", alpha=0.0027, phase = 1, method = "sw")
```

```{r}
# out of control points

print(paste("Total out of control data points : ", length(which(T2.2.3$t2 > T2.2.3$ucl)))) 
print(paste("Out of control data points: ", list(which(T2.2.3$t2 > T2.2.3$ucl))))

# total data points

print(paste("Total remaining data points", dim(df.pca)[1]))
```


# m-CUSUM charts for Phase 1 Analysis: 2nd Iteration

## Cycle: 1
```{r}
mcusum.2.1 <- mult.chart(df.pca, type = "mcusum2", alpha=0.0027, k = 1.5, h = 6, method = "sw")
```

```{r}
# out of control points

print(paste("Total out of control data points : ", length(which(mcusum.2.1$t2 > mcusum.2.1$ucl)))) 
print(paste("Out of control data points: ", list(which(mcusum.2.1$t2 > mcusum.2.1$ucl))))

# total data points

print(paste("Total remaining data points", dim(df.pca)[1]))
```


# Conclusion

```{r}
print(paste("Total data points in original data: ", dim(df)[1]))
print(paste("Total data points after removing out of control data points : ", dim(df.pca)[1]))
print(paste("Total out of control data points : ", dim(df)[1] - dim(df.pca)[1]))

```

## Phase II Analysis parameters:

Using remaining data points we can set up a monitoring scheme for future missions. The Hotelling T2 parameters for Phase II Analysis are:

```{r}
T2 <- mult.chart(df.pca, type = "t2", alpha=0.0027, method = "sw")

Xmv <- T2$Xmv
print("Means of Principal Components:")
Xmv

S <- T2$covariance
print("Covariance of Principal Components:")
S
```












